---
title: "Importing a Dataset and Fitting Models"
author: "George Kachergis"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
        fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Importing a Dataset and Fitting Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Importing a Dataset

Cross-situational word learning experiments are assumed to have a training phase, and a testing phase.
For the training phase, participants (and models) are presented with a subset of words and referents (i.e. objects) on each trial.
To allow for designs that present an unequal number of words and objects on each trial, we use a list format.

If you have data from an experiment that you would like to fit models to, or if you have 

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```



```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:


