---
title: "A large-scale comparison of several cross-situational word learning models"
bibliography: refs.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf George Kachergis (kachergis@stanford.edu)} \\ Department of Psychology, 450 Jane Stanford Way \\ Stanford, CA 94305 USA
    \AND {\large \bf Michael C.~Frank (mcfrank@stanford.edu)} \\ Department of Psychology, 450 Jane Stanford Way \\ Stanford, CA 94305 USA}

abstract: >
    How people learn word meanings from ambiguous situations has been a 
    frequent target of empirical work as well as numerous computational models, 
    and yet there is little consensus on the main mechanisms underpinning cross-
    situational word learning. The lack of consensus is in part due to the large 
    number of disparate studies and models, and lack of systematic model comparisons
    across a wide range of studies. This study seeks to remedy this, comparing the 
    performance of several extant cross-situational word learning models on a 
    dataset consisting of 1,696 participants spread across 44 experimental conditions.
    Using cross-validated parameters, we fit multiple models representing theories of both
    associative learning and hypothesis-testing theories of word learning, and discuss
    which models mimic each other, and which mechanisms are overall best able to account 
    for the data. Finally, we test the models' best-fitting parameters ability to generalize
    to a set of additional experiments, including developmental data.
    
keywords: >
    cross-situational word learning; model comparison; language learning
    
output: cogsci2016::cogsci_paper
#final-submission: \cogscifinalcopy
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, 
                      message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
```

# Introduction




# First-Level Headings

First level headings should be in 12 point , initial caps, bold and
centered. Leave one line space above the heading and 1/4~line space
below the heading.


## Two-column images


You might want to display a wide figure across both columns. To do this, you change the `fig.env` chunk option to `figure*`. To align the image in the center of the page, set `fig.align` option to `center`. To format the width of your caption text, you set the `num.cols.cap` option to `2`.

```{r 2-col-image, fig.env = "figure*", fig.pos = "h", fig.width=4, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns. And the caption text is limited to 0.8 of the width of the document."}
img <- png::readPNG("figs/walrus.png")
grid::grid.raster(img)
```

## One-column images

Single column is the default option, but if you want set it explicitly, set `fig.env` to `figure`. Notice that the `num.cols` option for the caption width is set to `1`.

```{r image, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "One column image."}
img <- png::readPNG("figs/lab_logo_stanford.png")
grid::grid.raster(img)
```


```{r plot, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2, fig.height=2, fig.cap = "R plot" }
x <- 0:100
y <- 2 * (x + rnorm(length(x), sd = 3) + 3)

ggplot2::ggplot(data = data.frame(x, y), 
                aes(x = x, y = y)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```


## Tables


```{r xtable, results="asis"}
n <- 100
x <- rnorm(n)
y <- 2*x + rnorm(n)
out <- lm(y ~ x)

tab1 <- xtable::xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2), 
                       caption = "This table prints across one column.")

print(tab1, type="latex", comment = F, table.placement = "H")
```

# Acknowledgements

Thanks to Chen Yu for allowing us to include data from several experimental conditions.

# References 

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent
